### 2022-07-26 src/text/0-abstract.tex
Add latexindent pre commit hook
Run tex indentation
Use shortcuts more consistently

\abstract{
	% TODO: the feature are still to detailed and partly copy and pasted form the ICRS abstract...avoid this
	% context heading (optional)
	% {} leave it empty if necessary  
	%{Gammapy context:}
	Historically the data as well as analysis software in gamma-ray astronomy is
	proprietary to the experiments. With the future Cherenkov Telescope Array
	(CTA), which will be operated as an open gamma-ray observatory with public
	data, there is a corresponding need for open high-level analysis software. In
	this article we present the first major version v1.0 of \gammapy, a
	community-developed open-source Python package for gamma-ray astronomy.
	 We present its general design and provide an%
	overview of the analysis methods and features it implements. Starting from
	event lists and a description of the specific instrument response functions
	(IRF) stored in open FITS based data formats, \gammapy implements . Thereby it
	handles the dependency of the IRFs with time, energy as well as position on the
	sky. It offers a variety of background estimation methods for spectral, spatial
	and spectro-morphological analysis. Counts, background and IRFs data are
	bundled in datasets and can be serialised, rebinned and stacked.
	 \gammapy supports to model binned data using%
	Poisson maximum likelihood fitting. It comes with built-in spectral, spatial
	and temporal models as well as support for custom user models, to model e.g.
	energy dependent morphology of gamma-ray sources. Multiple datasets can be
	combined in a joint-likelihood approach to either handle time dependent IRFs,
	different classes of events or combination of data from multiple instruments.
	Gammapy also implements methods to estimate flux points, including likelihood
	profiles per energy bin, light curves as well as flux and signficance maps in
	energy bins.  We further describe the general%
	development approach and how \gammapy integrates into ecosystem of other
	scientific and astronomical Python packages. We also present analysis examples
	with simulated CTA data and provide results of scientific validation analyses
	using data of existing instruments such as \hess and \fermi.

	% The gamma-ray can be considered as the last frontier 
	% In the last two decades the measurement of gamma-ray emission from the universe has 
	% evolved from a niche

>>>>

\abstract{
	% TODO: the feature are still to detailed and partly copy and pasted form the ICRS abstract...avoid this
	% context heading (optional)
	% {} leave it empty if necessary
	%{Gammapy context:}
	Historically the data as well as analysis software in \gammaray astronomy is
	proprietary to the experiments. With the future Cherenkov Telescope Array
	(CTA), which will be operated as an open \gammaray observatory with public
	data, there is a corresponding need for open high-level analysis software. In
	this article we present the first major version v1.0 of \gammapy, a
	community-developed open-source Python package for \gammaray astronomy. We
	present its general design and provide an%
	overview of the analysis methods and features it implements. Starting from
	event lists and a description of the specific instrument response functions
	(IRF) stored in open FITS based data formats, \gammapy implements . Thereby it
	handles the dependency of the IRFs with time, energy as well as position on the
	sky. It offers a variety of background estimation methods for spectral, spatial
	and spectro-morphological analysis. Counts, background and IRFs data are
	bundled in datasets and can be serialised, rebinned and stacked. \gammapy
	supports to model binned data using Poisson%
	maximum likelihood fitting. It comes with built-in spectral, spatial and
	temporal models as well as support for custom user models, to model e.g.,
	energy dependent morphology of \gammaray sources. Multiple datasets can be
	combined in a joint-likelihood approach to either handle time dependent IRFs,
	different classes of events or combination of data from multiple instruments.
	Gammapy also implements methods to estimate flux points, including likelihood
	profiles per energy bin, light curves as well as flux and signficance maps in
	energy bins.  We further describe the general%
	development approach and how \gammapy integrates into ecosystem of other
	scientific and astronomical Python packages. We also present analysis examples
	with simulated CTA data and provide results of scientific validation analyses
	using data of existing instruments such as \hess and \fermi.

	% The \gammaray can be considered as the last frontier
	% In the last two decades the measurement of \gammaray emission from the universe has
	% evolved from a niche

### 2022-07-26 src/text/1-introduction.tex
Use shortcuts more consistently

%% Gamma-ray astronomy
Gamma-ray astronomy is a rather young field of research. The \gammaray~
range of the electromagnetic spectrum provides us with insight to the
most energetic process in the universe such as surroundings of
black holes and remnants of supernova explosions. As in other
branches of astronomy \gammarays can be observed by both
satellite as well as ground based instruments.
Very high energy cosmic gamma-rays reaching Earth interact in the atmosphere and
create a large shower of secondary particles that can be observed from the ground.
Ground-based gamma-ray astronomy relies on this phenomenon to detect the
primary gamma-ray photons and measure their direction and energy.
It covers the energy range from fews tens of GeV up to the PeV.
There are two main categories of instrument \citep{2015CRPhy..16..610D}.
Imaging Atmospheric Cherenkov Telescopes (IACT) make images of atmospheric showers
by detecting the Cherenkov radiation emitted by the cascading charged particles and
use these images to reconstruct the properties of the incident particle.
Those instruments have a limited field of view (FoV) and duty cycle, but
good energy and angular resolution.

\begin{figure*}[t]
	\centering
	\includegraphics[height=\textwidth, angle=270]{static/gammapy-big-picture}
    \caption{ Gammapy is a Python package
		for high-level gamma-ray data analysis. Using event lists, exposures and point
		spread functions as input you can use it to generate science results such as
		images, spectra, light curves or source catalogs. So far it has been used to
		simulate and analyse H.E.S.S., CTA,  and \textit{Fermi}-LAT data. As demonstrated
		by several studies, it will also be applied to e.g., VERITAS, MAGIC or HAWC data
		in the future. }
	\label{fig:big-picture}

@@

%% Context
Ground based gamma-ray astronomy has been historically structured
by experiments run by independent collaborations relying
on their own proprietary data and analysis software developed as part of the
instrument. While this model has been very successful so far, it does not
permit easy combination of data from several instruments and is therefore
a limitation for interoperability of existing facilities. Especially
because the different detection techniques have complementary
properties.

@@

%% Context : formats
In practice, the data-reduction workflow of all gamma-ray observatories
is remarkably similar.
After data calibration, shower events are reconstructed and
gamma/hadron separation is applied to build lists of gamma-ray like events.
The latter are then used to derive scientific results, such as spectra, sky maps
or light curves, taking into account the specific instrument response functions (IRF).
The information in this high data level is independent on
the data reduction, and eventually of the detection technique. This implies,
for example, that data from IACT and particle samplers can be represented
within the same model. The efforts to prototype a format usable by various instruments
converged in the so-called \textit{Data Format for Gamma-ray Astronomy}
initiative~\citep{gadf_proc, gadf_universe}, abbreviated in
\texttt{gamma-astro-data-format} (GADF). The latter proposes prototypical
specifications to produce files based on the flexible image transport system
(FITS) format~\citep{fits} encapsulating this high-level information. This is
realized by storing a list of gamma-like events with their measured quantites
(energy, direction, arrival time) and a parametrisation of the response of the
system. (see Sec.~\ref{ssec:gammapy-data} and Sec.~\ref{ssec:gammapy-irf} for
more information).

@@

%% Gammapy: concept and goals
%%% Question: starting date? Is is relevant, what is the real start?
The Gammapy project started in 2015 with the objective of building a flexible and
efficient software library for very high energy gamma-ray data analysis \citep{gammapy_2015}.
This is possible thanks to the definition of a common data format such as GADF, and thanks to
an open development with a broad community of contributors.

@@

%% Gammapy: from first developments to validation and selection
The H.E.S.S. collaboration released a limited test dataset (about 50 hours of
observations taken between 2004 and 2008) based  on the GADF DL3 format \citep{HESS_DR1}.
This data release served as a basis for analysis tools validation \cite[see e.g.]{Mohrmann2019}.

@@

%% Paper outline
In this article, we describe the general structure of the Gammapy package and
its main concepts. In Section \ref{sec:analysis-workflow-overview}, we present
the data analysis workflow in very high energy gamma-ray astronomy. Then we
show how this workflow is reflected in the structure of the Gammapy package and
describe the various subpackages it contains. Section \ref{sec:applications}
presents a number of applications, while Section \ref{sec:gammapy-project}
discusses the project organization.

>>>>

%% \gammaray astronomy
\gammaray astronomy is a rather young field of research. The \gammaray~
range of the electromagnetic spectrum provides us with insight to the
most energetic process in the universe such as surroundings of
black holes and remnants of supernova explosions. As in other
branches of astronomy \gammarays can be observed by both
satellite as well as ground based instruments.
Very high energy cosmic \gammarays reaching Earth interact in the atmosphere and
create a large shower of secondary particles that can be observed from the ground.
Ground-based \gammaray astronomy relies on this phenomenon to detect the
primary \gammaray photons and measure their direction and energy.
It covers the energy range from fews tens of GeV up to the PeV.
There are two main categories of instrument \citep{2015CRPhy..16..610D}.
Imaging Atmospheric Cherenkov Telescopes (IACT) make images of atmospheric showers
by detecting the Cherenkov radiation emitted by the cascading charged particles and
use these images to reconstruct the properties of the incident particle.
Those instruments have a limited field of view (FoV) and duty cycle, but
good energy and angular resolution.

\begin{figure*}[t]
	\centering
	\includegraphics[height=\textwidth, angle=270]{static/gammapy-big-picture}
    \caption{ Gammapy is a Python package
		for high-level \gammaray data analysis. Using event lists, exposures and point
		spread functions as input you can use it to generate science results such as
		images, spectra, light curves or source catalogs. So far it has been used to
		simulate and analyse \hess, \cta,  and \fermi data. As demonstrated
		by several studies, it will also be applied to e.g., \veritas, \magic or \hawc data
		in the future. }
	\label{fig:big-picture}

@@

%% Context
Ground based \gammaray astronomy has been historically structured
by experiments run by independent collaborations relying
on their own proprietary data and analysis software developed as part of the
instrument. While this model has been very successful so far, it does not
permit easy combination of data from several instruments and is therefore
a limitation for interoperability of existing facilities. Especially
because the different detection techniques have complementary
properties.

@@

%% Context : formats
In practice, the data-reduction workflow of all \gammaray observatories
is remarkably similar.
After data calibration, shower events are reconstructed and
gamma/hadron separation is applied to build lists of \gammaray like events.
The latter are then used to derive scientific results, such as spectra, sky maps
or light curves, taking into account the specific instrument response functions (IRF).
The information in this high data level is independent on
the data reduction, and eventually of the detection technique. This implies,
for example, that data from IACT and particle samplers can be represented
within the same model. The efforts to prototype a format usable by various instruments
converged in the so-called \textit{Data Format for \gammaray Astronomy}
initiative~\citep{gadf_proc, gadf_universe}, abbreviated in
\texttt{gamma-astro-data-format} (\gadf). The latter proposes prototypical
specifications to produce files based on the flexible image transport system
(FITS) format~\citep{fits} encapsulating this high-level information. This is
realized by storing a list of gamma-like events with their measured quantites
(energy, direction, arrival time) and a parametrisation of the response of the
system. (see Sec.~\ref{ssec:gammapy-data} and Sec.~\ref{ssec:gammapy-irf} for
more information).

@@

%% Gammapy: concept and goals
%%% Question: starting date? Is is relevant, what is the real start?
The Gammapy project started in 2015 with the objective of building a flexible and
efficient software library for very high energy \gammaray data analysis \citep{gammapy_2015}.
This is possible thanks to the definition of a common data format such as \gadf, and thanks to
an open development with a broad community of contributors.

@@

%% Gammapy: from first developments to validation and selection
The H.E.S.S. collaboration released a limited test dataset (about 50 hours of
observations taken between 2004 and 2008) based  on the \gadf DL3 format \citep{HESS_DR1}.
This data release served as a basis for analysis tools validation \cite[see e.g.]{Mohrmann2019}.

@@

%% Paper outline
In this article, we describe the general structure of the Gammapy package and
its main concepts. In Section \ref{sec:analysis-workflow-overview}, we present
the data analysis workflow in very high energy \gammaray astronomy. Then we
show how this workflow is reflected in the structure of the Gammapy package and
describe the various subpackages it contains. Section \ref{sec:applications}
presents a number of applications, while Section \ref{sec:gammapy-project}
discusses the project organization.

### 2022-07-26 src/text/2-package-subsections/stats.tex
Minor fixes
Use shortcuts more consistently

\subsection{gammapy.stats}
\label{ssec:gammapy-stats}
The \code{gammapy.stats} subpackage contains the fit statistics and associated
statistical estimators that are commonly used in gamma-ray astronomy. In
general, gamma-ray observations count Poisson-distributed events at various sky
positions, and contain both signal and background events. Estimation of the
number of signal events is done through likelihood maximization. In Gammapy,
the fit statistics are Poisson log-likelihood functions normalized like
chi-squares, i.e., they follow the expression $2 \times log L$, where $L$ is
the likelihood function used. The statistic function used when the expected
number of background events is known is the \emph{Cash} statistic
~\citep{Cash}. It is used by datasets using background templates such as the
MapDataset. When the number of background events is unknown and an off
measurement where only background events are expected is used, the statistic
function is WStat. It is a profile log-likelihood statistic where background
counts are marginalized parameters. It is used by datasets containing off
counts measurements such as the SpectrumDatasetOnOff, used for classical
spectral analysis.

>>>>

\subsection{gammapy.stats}
\label{ssec:gammapy-stats}
The \code{gammapy.stats} subpackage contains the fit statistics and associated
statistical estimators that are commonly used in \gammaray astronomy. In
general, \gammaray observations count Poisson-distributed events at various sky
positions, and contain both signal and background events. Estimation of the
number of signal events is done through likelihood maximization. In Gammapy,
the fit statistics are Poisson log-likelihood functions normalized like
chi-squares, i.e., they follow the expression $2 \times log L$, where $L$ is
the likelihood function used. The statistic function used when the expected
number of background events is known is the \emph{Cash} statistic
~\citep{Cash}. It is used by datasets using background templates such as the
\code{MapDataset}. When the number of background events is unknown and an off
measurement where only background events are expected is used, the statistic
function is WStat. It is a profile log-likelihood statistic where background
counts are marginalized parameters. It is used by datasets containing off
counts measurements such as the SpectrumDatasetOnOff, used for classical
spectral analysis.

### 2022-07-26 src/text/2-package.tex
Use shortcuts more consistently

\section{Gamma-ray Data Analysis}
\label{sec:gamma-ray-data-analysis}

@@

The other sequence, referred to as high-level analysis, deals with the extraction of physical
quantities related to gamma-ray sources and the production of high-level products such as spectra,
lightcurves and catalogs. The methods and tools applied here are more generic and are broadly
shared across the field. They also frequently imply joint analysis of multi-instrument data.
To extract physically relevant information, the measured data are usually compared to a
model of the expected gamma-ray emitters in the instrument field-of-view using statistical
techniques such as maximum likelihood.

@@

\begin{align}
   R(p, E|p_{\rm true}, E_{\rm true}) = &A_{\rm eff}(p_{\rm true}, E_{\rm true}) \times\\
    &PSF(p|p_{\rm true}, E_{\rm true}) \times\\
    &E_{\rm disp}(E|p_{\rm true}, E_{\rm true})
\end{align}
where:
\begin{itemize}
\item $A_{\rm eff}(p_{\rm true}, E_{\rm true})$ is the effective collection area of the detector. It is the product
  of the detector collection area times its detection efficiency at true energy $E_{\rm true}$ and position $p_{\rm true}$.
\item $PSF(p|p_{\rm true}, E_{\rm true})$ is the point spread function. It gives the probability of
  measuring a direction $p$ when the true direction is $p_{\rm true}$ and the true energy is $E_{\rm true}$.
  Gamma-ray instruments consider the probability density of the angular separation between true and reconstructed directions
  $\delta p = p_{\rm true} - p$, i.e. $PSF(\delta p|p_{\rm true}, E_{\rm true})$.
\item $E_{\rm disp}(E|p_{\rm true}, E_{\rm true})$ is the energy dispersion. It gives the probability to
  reconstruct the photon at energy $E$ when the true energy is $E_{\rm true}$ and the true position :$p_{\rm true}$.
  Gamma-ray instruments consider the probability density of the migration $\mu=\frac{E}{E_{\rm true}}$,
  i.e. $E_{\rm disp}(\mu|p_{\rm true}, E_{\rm true})$.
\end{itemize}

Gamma-ray data at the Data Level 3 therefore consists in lists of gamma-like events and their
corresponding instrument response functions (IRFs). The latter include the aforementioned
effective area, point spread function (PSF), energy dispersion and residual hadronic background.
The handling of DL3 data is performed by classes and methods
in the gammapy.data  (see \ref{ssec:gammapy-data}) and the gammapy.irf
(see \ref{ssec:gammapy-irf}) subpackages.

@@

The next step is then typically to model and fit the datasets, either
individually, or in a joint likelihood analysis. For this purpose Gammapy
provides a uniform interface to multiple fitting backends. It also provides
a variety of :ref:`built in models <model-gallery>`. This includes spectral,
spatial and temporal model classes to describe the gamma-ray emission in the sky.
Where spectral models can be simple analytical models or more complex ones from radiation
mechanisms of accelerated particle populations (e.g. inverse Compton or $\pi^{o}$ decay).
Independently or subsequently to the global modelling, the data can be
re-grouped to compute flux points, light curves and flux as well as significance
maps in energy bands.

>>>>

\section{\gammaray Data Analysis}
\label{sec:gammaray-data-analysis}

@@

The other sequence, referred to as high-level analysis, deals with the extraction of physical
quantities related to \gammaray sources and the production of high-level products such as spectra,
lightcurves and catalogs. The methods and tools applied here are more generic and are broadly
shared across the field. They also frequently imply joint analysis of multi-instrument data.
To extract physically relevant information, the measured data are usually compared to a
model of the expected \gammaray emitters in the instrument field-of-view using statistical
techniques such as maximum likelihood.

@@

\begin{align}
   R(p, E|p_{\rm true}, E_{\rm true}) = &A_{\rm eff}(p_{\rm true}, E_{\rm true}) \times\\
    &PSF(p|p_{\rm true}, E_{\rm true}) \times\\
    &E_{\rm disp}(E|p_{\rm true}, E_{\rm true})
\end{align}
where:
\begin{itemize}
\item $A_{\rm eff}(p_{\rm true}, E_{\rm true})$ is the effective collection area of the detector. It is the product
  of the detector collection area times its detection efficiency at true energy $E_{\rm true}$ and position $p_{\rm true}$.
\item $PSF(p|p_{\rm true}, E_{\rm true})$ is the point spread function. It gives the probability of
  measuring a direction $p$ when the true direction is $p_{\rm true}$ and the true energy is $E_{\rm true}$.
  \gammaray instruments consider the probability density of the angular separation between true and reconstructed directions
  $\delta p = p_{\rm true} - p$, i.e. $PSF(\delta p|p_{\rm true}, E_{\rm true})$.
\item $E_{\rm disp}(E|p_{\rm true}, E_{\rm true})$ is the energy dispersion. It gives the probability to
  reconstruct the photon at energy $E$ when the true energy is $E_{\rm true}$ and the true position :$p_{\rm true}$.
  \gammaray instruments consider the probability density of the migration $\mu=\frac{E}{E_{\rm true}}$,
  i.e. $E_{\rm disp}(\mu|p_{\rm true}, E_{\rm true})$.
\end{itemize}

\gammaray data at the Data Level 3 therefore consists in lists of gamma-like events and their
corresponding instrument response functions (IRFs). The latter include the aforementioned
effective area, point spread function (PSF), energy dispersion and residual hadronic background.
The handling of DL3 data is performed by classes and methods
in the gammapy.data  (see \ref{ssec:gammapy-data}) and the gammapy.irf
(see \ref{ssec:gammapy-irf}) subpackages.

@@

The next step is then typically to model and fit the datasets, either
individually, or in a joint likelihood analysis. For this purpose Gammapy
provides a uniform interface to multiple fitting backends. It also provides
a variety of :ref:`built in models <model-gallery>`. This includes spectral,
spatial and temporal model classes to describe the \gammaray emission in the sky.
Where spectral models can be simple analytical models or more complex ones from radiation
mechanisms of accelerated particle populations (e.g. inverse Compton or $\pi^{o}$ decay).
Independently or subsequently to the global modelling, the data can be
re-grouped to compute flux points, light curves and flux as well as significance
maps in energy bands.

### 2022-07-26 src/text/3-applications-subsections/1D-analysis.tex
Use shortcuts more consistently

\subsection{1D Analysis}
\label{ssec:1d-analysis}
One of the most common analysis cases in gamma-ray astronomy is measuring the
spectrum of a source in a given region defined on the sky, in conventional
astronomy also called aperture photometry. The spectrum is typically measured
in two steps: first a parametric spectral model is fitted to the data and
secondly flux points are computed in a pre-defined set of energy bins. The
result of such an analysis performed on three simulated CTA observations is
shown in Fig.~\ref{fig:cta_galactic_center}. In this case the spectrum was
measured in a circular aperture centered on the Galactic Center, in
\gammaray~astronomy often called "on region". For such analysis the users first
chooses a region of interest and energy binning, both defined by a
\code{RegionGeom}. In a second step the events and instrument response are binned
into maps of this geometry, by the \code{SpectrumDatasetMaker}. All the data and
reduced instrument response are bundled into a \code{SpectrumDataset}. To estimate
the expected background in the on region a "reflected regions" background
method was used~\cite{Berge07}, represented in \gammapy by the
\code{ReflectedRegionsBackgroundMaker} class. The resulting reflected regions are
illustrated for all three observations on top of the map of counts. After
reduction of the data it was modelled using a forward-folding method and a
power-law spectral shape, using the \code{PowerLawSpectralModel} and \code{Fit} class.
Based on this best fit model the final flux points and corresponding
log-likelihood profiles are computed using the \code{FluxPointsEstimator}.

>>>>

\subsection{1D Analysis}
\label{ssec:1d-analysis}
One of the most common analysis cases in \gammaray astronomy is measuring the
spectrum of a source in a given region defined on the sky, in conventional
astronomy also called aperture photometry. The spectrum is typically measured
in two steps: first a parametric spectral model is fitted to the data and
secondly flux points are computed in a pre-defined set of energy bins. The
result of such an analysis performed on three simulated CTA observations is
shown in Fig.~\ref{fig:cta_galactic_center}. In this case the spectrum was
measured in a circular aperture centered on the Galactic Center, in
\gammaray~astronomy often called "on region". For such analysis the users first
chooses a region of interest and energy binning, both defined by a
\code{RegionGeom}. In a second step the events and instrument response are binned
into maps of this geometry, by the \code{SpectrumDatasetMaker}. All the data and
reduced instrument response are bundled into a \code{SpectrumDataset}. To estimate
the expected background in the on region a "reflected regions" background
method was used~\cite{Berge07}, represented in \gammapy by the
\code{ReflectedRegionsBackgroundMaker} class. The resulting reflected regions are
illustrated for all three observations on top of the map of counts. After
reduction of the data it was modelled using a forward-folding method and a
power-law spectral shape, using the \code{PowerLawSpectralModel} and \code{Fit} class.
Based on this best fit model the final flux points and corresponding
log-likelihood profiles are computed using the \code{FluxPointsEstimator}.

### 2022-07-26 src/text/3-applications-subsections/population-studies.tex
Use shortcuts more consistently

Sky surveys have a large potential for new source detections, and new phenomena
discovery. They also offer less selection bias to perform source population
studies over a large set of coherently detected and modelled objects. Early
versions of \textit{gammapy} were developed in parallel of the preparation of
the HESS Galactic plane survey catalog~\citep[HGPS][]{2018A&A...612A...1H} and
the associated PWN and SNR populations studies~\citep{2018A&A...612A...2H,
	2018A&A...612A...3H}.

The increase in sensitivity  and resolution provided by new generation of
instruments scale up the number of sources detectable and the complexity of the
models needed to represent them accurately. As an example If we compare the
results of the HGPS to the expectations from the CTA Galactic Plane survey
simulations we jump from 78 sources detected by HESS to about 500 detectable by
CTA~\citep{2021arXiv210903729R}.

@@

In short the production of catalogs from gamma-ray surveys can be divided in
four main steps : data-reduction; object detection; model fitting and model
selection; associations and classification. The IACTs data-reduction step is
done in the same way than described in the previous sections but scale-up to
few thousand of observations. The object detection step consists a minima in
finding local maxima in the significance, or TS maps,  given by the
ExcessMapEstimator, or TSMapEstimator, respectively.  Further refinements can
include for example  filtering and detection on these maps with techniques from
scikit-image package~\citep{scikit-image}, and outlier detection from
scikit-learn package~\citep{scikit-learn}. Tests on simulations shown that it
reduces the spurious detections at this stage and then speed up the next step
as less objects will have to be fitted simultaneously. During the modelling
step each object is alternatively fitted with different models in order to
determine their optimal parameters, and the best-candidate model. The
subpackage gammapy.modeling.models offers a large variety of choice, and the
possibility to add custom models.  Several spatial models (point-source, disk,
gaussian...), and spectral models (power-law, log-parabola...) may be tested
for each object, so the complexity of the problem increases rapidly in regions
crowded with multiple extended sources. Finally an object is discarded if its
best-fit model is not significantly preferred over the null hypothesis (no
source) comparing the difference in log-Likelihood between these two
hypotheses. For the association and classification step, that is tightly
connected to the population studies, we can a minima compare the fitted models
to the set of gammapy-ray catalogs available in gammapy.catalog. However
further multi-wavelength cross-matches are usually required to characterize the
sources.

>>>>

Sky surveys have a large potential for new source detections, and new phenomena
discovery. They also offer less selection bias to perform source population
studies over a large set of coherently detected and modelled objects. Early
versions of \textit{gammapy} were developed in parallel of the preparation of
the \hess Galactic plane survey catalog~\citep[HGPS][]{2018A&A...612A...1H} and
the associated PWN and SNR populations studies~\citep{2018A&A...612A...2H,
	2018A&A...612A...3H}.

The increase in sensitivity  and resolution provided by new generation of
instruments scale up the number of sources detectable and the complexity of the
models needed to represent them accurately. As an example If we compare the
results of the HGPS to the expectations from the \cta Galactic Plane survey
simulations we jump from 78 sources detected by \hess to about 500 detectable by
CTA~\citep{2021arXiv210903729R}.

@@

In short the production of catalogs from \gammaray surveys can be divided in
four main steps : data-reduction; object detection; model fitting and model
selection; associations and classification. The IACTs data-reduction step is
done in the same way than described in the previous sections but scale-up to
few thousand of observations. The object detection step consists a minima in
finding local maxima in the significance, or TS maps,  given by the
ExcessMapEstimator, or TSMapEstimator, respectively.  Further refinements can
include for example  filtering and detection on these maps with techniques from
scikit-image package~\citep{scikit-image}, and outlier detection from
scikit-learn package~\citep{scikit-learn}. Tests on simulations shown that it
reduces the spurious detections at this stage and then speed up the next step
as less objects will have to be fitted simultaneously. During the modelling
step each object is alternatively fitted with different models in order to
determine their optimal parameters, and the best-candidate model. The
subpackage gammapy.modeling.models offers a large variety of choice, and the
possibility to add custom models.  Several spatial models (point-source, disk,
gaussian...), and spectral models (power-law, log-parabola...) may be tested
for each object, so the complexity of the problem increases rapidly in regions
crowded with multiple extended sources. Finally an object is discarded if its
best-fit model is not significantly preferred over the null hypothesis (no
source) comparing the difference in log-Likelihood between these two
hypotheses. For the association and classification step, that is tightly
connected to the population studies, we can a minima compare the fitted models
to the set of gammapy-ray catalogs available in gammapy.catalog. However
further multi-wavelength cross-matches are usually required to characterize the
sources.

### 2022-07-26 src/text/4-project.tex
Use shortcuts more consistently

Gammapy is indexed in Astronomy Source Code
Library\footnote{\url{https://ascl.net/1711.014}} and
Zenodo\footnote{\url{https://zenodo.org/record/5721467o}} digital libraries for
software. The Zenodo record is synchronised with the codebase GitHub repository
so that every release triggers the update of the versioned record. In addition,
the next release of Gammapy will be added to the Open-source scientific
Software and Service Repository\footnote{\url{https://projectescape.eu/ossr}}
and indexed in the European Open Science Cloud
catalog\footnote{\url{https://eosc-portal.eu}}.

@@

\subsection{Release Cycle, Versioning, and Long-term Support}
With the first long term support (LTS) release v1.0, the \gammapy project
enters a new development phase. The development will change from
quick feature driven development to more stable maintenance
and user support driven developement. After v1.0 we foresee
a developement cycle with major, minor and bugfix releases.
We basically followg the development cycle of the Astropy
project. Thus we expect a major LTS release approximately
every two years, minor releases are planned every 6~months,
while bug-fix releases will happen as needed. While
bug-fix releases will not introduce API breaking changes,
we will work with a deprecation system for minor releases.
API breaking changes will be announced to user by runtime
warnings first and then implemented in the subsequent
minor release. This approach we consider a fair
compromise between the interests of users in a stable
package and the interest of developers to improve
and develop \gammapy in future. The developemnt cycle is described
in more detail in PIG 24 \todo{reference}.

>>>>

\gammapy is indexed in Astronomy Source Code
Library\footnote{\url{https://ascl.net/1711.014}} and
Zenodo\footnote{\url{https://zenodo.org/record/5721467o}} digital libraries for
software. The Zenodo record is synchronised with the codebase GitHub repository
so that every release triggers the update of the versioned record. In addition,
the next release of Gammapy will be added to the Open-source scientific
Software and Service Repository\footnote{\url{https://projectescape.eu/ossr}}
and indexed in the European Open Science Cloud
catalog\footnote{\url{https://eosc-portal.eu}}.

@@

\subsection{Release Cycle, Versioning, and Long-term Support}
\label{ssec:release-cycle}
With the first long term support (LTS) release v1.0, the \gammapy project
enters a new development phase. The development will change from
quick feature driven development to more stable maintenance
and user support driven developement. After v1.0 we foresee
a developement cycle with major, minor and bugfix releases.
We basically followg the development cycle of the Astropy
project. Thus we expect a major LTS release approximately
every two years, minor releases are planned every 6~months,
while bug-fix releases will happen as needed. While
bug-fix releases will not introduce API breaking changes,
we will work with a deprecation system for minor releases.
API breaking changes will be announced to user by runtime
warnings first and then implemented in the subsequent
minor release. This approach we consider a fair
compromise between the interests of users in a stable
package and the interest of developers to improve
and develop \gammapy in future. The developemnt cycle is described
in more detail in PIG 24 \todo{reference}.
